Semi-permeable vesicles: mathematical formulation, numerics and physics

### 2020-03-13 vesicle_code/docs/permeableVes/main.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Vesicle Area}
Since the vesicle merrane is semi-permeable, the vesicle's area
fluctuates.  Parameterizing the vesicle with $\xx$, its area is
\begin{align}
  A(t) = \frac{1}{2}\int_{\gamma} (\xx \cdot \nn)\, ds .
\end{align}
The time derivative of $A(t)$ includes terms due to the time derivative
of $\xx \cdot \nn$, and a term due to interface
stretching~\cite{lai-tse-hua2008}. However, this second term is zero
because of the local inextensibility condition. Therefore, 
\begin{align}
  \dot{A}(t) &= \frac{1}{2} \int_{\gamma} \pderiv{}{t} 
    (\xx \cdot \nn) \, ds = 
  \frac{1}{2} \int_{\gamma} (\dot{\xx} \cdot \nn)\, ds  + 
  \frac{1}{2} \int_{\gamma} (\xx \cdot \dot{\nn})\, ds.
\end{align}
The second integrand can be computed as
\begin{align}
  \int_\gamma (\xx \cdot \dot{\nn})\, ds 
    = -\int_\gamma \left(\xx^\perp \cdot \dot{\xx}_s\right)\, ds 
    = \int_\gamma \left(\xx_s^\perp \cdot \dot{\xx}\right)\, ds 
    = \int_\gamma (\dot{\xx} \cdot \nn)\, ds.
\end{align}
Therefore,
\begin{align}
  \dot{A}(t) &= \int_\gamma (\dot{\xx} \cdot \nn)\, ds 
  = \int_\gamma \left(\beta (\ff \cdot \nn)\nn + \SS[\ff]\right) 
      \cdot \nn\, ds 
  = \beta \int_\gamma (\ff \cdot \nn)\, ds,
\end{align}
where the last equality comes follows from the divergence theorem.
Finally, using integration by parts and the Frenet-Serret identities, we
have
\begin{align*}
  \dot{A}(t) &= \beta \int_\gamma \left( -\kappa_b \xx_{ssss} + 
    (\sigma \xx_s)_s \right) \cdot \nn \, ds \\
  &= \beta \int_\gamma \left(-\kappa_b \xx_{sss} + \sigma \xx_s 
    \right) \cdot \kappa \xx_s \, ds \\
  &= \beta \int_\gamma \left(\kappa_b \xx_{ss} \cdot 
    (\kappa \xx_s)_s \right) \, ds +
    \beta \int_\gamma \kappa \sigma \, ds \\
  &= \beta \int_\gamma \left(\kappa_b \kappa (\xx_{ss} \cdot \xx_{ss}) + 
    \kappa_b \kappa_s (\xx_{ss} \cdot \xx_s) \right) \, ds +
    \beta \int_\gamma \kappa \sigma \, ds \\
  &= \beta \int_\gamma \left(\kappa_b \kappa^3 + \kappa \sigma \right) 
    \, ds.
\end{align*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Numerical Methods}
\subsection{Discretization in Space}
We apply a collocation method where the vesicle is discretized at $N$
points. We need to compute derivatives of functions defined on the
vesicle boundary, and these are done with Fourier differentiation.
Finally, to compute the single-layer potential, we use an eight-order
Alpert quadrature rule~\cite{alp1999} which resolves the logarithmic
singularity of the integrand.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Vesicle Area}
Since the vesicle merrane is semi-permeable, the vesicle's area
fluctuates.  Parameterizing the vesicle with $\xx$, its area is
\begin{align}
  A(t) = \frac{1}{2}\int_{\gamma} (\xx \cdot \nn)\, ds .
\end{align}
The time derivative of $A(t)$ includes terms due to the time derivative
of $\xx \cdot \nn$, and a term due to interface
stretching~\cite{lai-tse-hua2008}. However, this second term is zero
because of the local inextensibility condition. Therefore, 
\begin{align}
  \dot{A}(t) &= \frac{1}{2} \int_{\gamma} \pderiv{}{t} 
    (\xx \cdot \nn) \, ds = 
  \frac{1}{2} \int_{\gamma} (\dot{\xx} \cdot \nn)\, ds  + 
  \frac{1}{2} \int_{\gamma} (\xx \cdot \dot{\nn})\, ds.
\end{align}
The second integrand can be computed as
\begin{align}
  \int_\gamma (\xx \cdot \dot{\nn})\, ds 
    = -\int_\gamma \left(\xx^\perp \cdot \dot{\xx}_s\right)\, ds 
    = \int_\gamma \left(\xx_s^\perp \cdot \dot{\xx}\right)\, ds 
    = \int_\gamma (\dot{\xx} \cdot \nn)\, ds.
\end{align}
Therefore,
\begin{align}
  \dot{A}(t) &= \int_\gamma (\dot{\xx} \cdot \nn)\, ds 
  = \int_\gamma \left(\beta (\ff \cdot \nn)\nn + \SS[\ff]\right) 
      \cdot \nn\, ds 
  = \beta \int_\gamma (\ff \cdot \nn)\, ds,
\end{align}
where the last equality comes follows from the divergence theorem.
Finally, using integration by parts and the Frenet-Serret identities, we
have
\begin{align*}
  \dot{A}(t) &= \beta \int_\gamma \left( -\kappa_b \xx_{ssss} + 
    (\sigma \xx_s)_s \right) \cdot \nn \, ds \\
  &= \beta \int_\gamma \left(-\kappa_b \xx_{sss} + \sigma \xx_s 
    \right) \cdot \kappa \xx_s \, ds \\
  &= \beta \int_\gamma \left(\kappa_b \xx_{ss} \cdot 
    (\kappa \xx_s)_s + \kappa \sigma \right) \, ds \\
  &= \beta \int_\gamma \left(\kappa_b \kappa (\xx_{ss} \cdot \xx_{ss}) + 
    \kappa_b \kappa_s (\xx_{ss} \cdot \xx_s) + \kappa \sigma \right) 
    \, ds \\
  &= \beta \int_\gamma \left(\kappa_b \kappa^3 + \kappa \sigma \right) 
    \, ds.
\end{align*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Numerical Methods}
\subsection{Discretization in Space}
We apply a collocation method where the vesicle is discretized at $N$
points. We need to compute derivatives of functions defined on the
vesicle boundary, and these are done with Fourier differentiation.
Finally, to compute the single-layer potential, we use an eighth-order
Alpert quadrature rule~\cite{alp1999} which resolves the logarithmic
singularity of the integrand.
